### 2020-10-08 serial.tex
abstract tweaks

The reproduction number \RR and the exponential growth rate $r$ are critical quantities for characterizing outbreak dynamics.
These quantities are linked by generation intervals---defined as the time between infection and transmission.
Because generation intervals are difficult to measure directly, serial intervals---defined as the onset of symptoms in a chain of transmission---are often substituted.
However, recent studies suggest that the two intervals give different estimates of \RR from $r$.
We show that \emph{forward-looking} serial intervals, which begin in a certain period, correctly describe the renewal process of symptomatic cases and therefore link \RR with $r$.
These forward intervals are different from \emph{intrinsic} intervals, which neglect population-level dynamics, and \emph{backward-looking} intervals, which end in a certain period.
Instead, the forward-looking intervals depend on $r$:
When an epidemic is growing exponentially, infectors that develop symptoms on a given day will have shorter incubation periods, on average, than their infectees, thereby affecting the serial-interval distribution.
We provide a heuristic method for addressing biases that can arise from neglecting changes in serial intervals. 
We apply the method to estimation of \RR for the COVID-19 outbreak in China, outside Hubei province, using serial-interval data collected between January 21--February 8, 2020. 
Our analysis shows that using incorrectly defined serial intervals can bias estimates of initial \RR by up to a factor of 2.6 in the context of COVID-19.
This study demonstrates the importance of early contact tracing efforts and provides a rationale for reassessing generation intervals, serial intervals, and \RR\ estimates, for COVID-19.

>>>>

The reproduction number \RR and the growth rate $r$ are critical quantities for characterizing outbreak dynamics.
They are linked by generation intervals, the time between infection and onward transmission.
Because generation intervals are difficult to observe, epidemiologists often substitute serial intervals, the time between symptom onset in successive links in a transmission chain.
Recent studies suggest that such substitution can bias estimates of \RR based on $r$.
Here we explore how these intervals vary over the course of an epidemic, and the implications for \RR estimation.
\emph{Forward-looking} serial intervals, measuring time forward from symptom onset of a focal individual, correctly describe the renewal process of symptomatic cases and therefore reliably link \RR with $r$.
In contrast, \emph{backward-looking} intervals, which measure time backward, and \emph{intrinsic} intervals, which neglect population-level dynamics, lead to incorrect \RR estimates.
Forward-looking intervals are affected both by epidemic dynamics and by censoring, changing in complex ways over the course of an epidemic.
We derive a heuristic method for addressing biases that arise from neglecting changes in serial intervals. 
We apply the method to early serial-interval-based estimates of  \RR for the COVID-19 outbreak in China outside Hubei province,
showing that using improperly defined serial intervals in this context can bias estimates of initial \RR by up to a factor of 2.6.
This study demonstrates the importance of early contact-tracing efforts and provides a framework for reassessing generation intervals, serial intervals, and \RR\ estimates for COVID-19.
